
```{r}
if(!require(plyr)){
  install.packages("plyr")
	library(plyr)
}

if(!require(dplyr)){
  install.packages("dplyr")
	library(dplyr)
}

if(!require(lubridate)){
  install.packages("lubridate")
	library(lubridate)
}

if(!require(ggplot2)){
  install.packages("ggplot2")
	library(ggplot2)
}
```
## Unzip the file and load files in the specific directory
```{r}
zip_file <- file.path("activity.zip")
unzip("activity.zip")
```
##Load data and process/transform the data as require
```{r}
activity <- read.csv("activity.csv",colClasses = c("numeric", "character","integer"))

dim(activity)
str(activity)
summary(activity)
names(activity)
```
# Part One
## mean total number of steps taken per day. Ignore the missing values in the dataset.
```{r}

total_steps <- tapply(activity$steps, activity$date, FUN = sum, na.rm = TRUE)
activity$date <- ymd(activity$date)

mean(total_steps)
median(total_steps)

```
## Calculate the total number of steps taken per day
```{r}
steps <- activity
steps <- filter(steps,!is.na(steps))
steps <- group_by(steps, date) 
steps <- summarize(steps, steps=sum(steps))

dim(steps)
class(steps)
str(steps)
summary(steps)
names(steps)
head(steps)
tail(steps)

```
## Make a histogram of the total number of steps taken each day
```{r, echo=FALSE}
ggplot(steps, aes(x=steps)) +
	geom_histogram(bins=20, colour="blue", fill="blue") +
	guides(fill=FALSE) +
	xlab("Steps")+ ylab("Count")+  
	labs(title= "Histogram of the total number of steps taken each day\n(NA removed)")
#dev.off()

```
# Part Two
## What is the average daily activity pattern?
```{r}
daily <- activity
daily <- filter(daily,!is.na(steps))
daily <- group_by(daily, interval) 
daily <- summarize(daily, steps=mean(steps))

dim(daily)
class(daily)
str(daily)
summary(daily)
names(daily)
head(daily)
tail(daily)
```
## time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, echo=FALSE}
plot(daily, type = "l",col="blue",xlab="5-minute interval", ylab="Average steps")
title("Average number of steps taken in 5-minutes intervals")
#dev.off()
```
## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
## [1] 835
```{r}
daily[which.max(daily$steps), ]$interval
```
## Imputing missing values
## the total number of missing values
```{r}
count(activity[complete.cases(activity[1,2,3]),])
count(na.omit(activity))
count(activity)
count(activity)-count(na.omit(activity))
sum(is.na(activity))
```
## Imputing missing values
##Create a new dataset where missing values in the dataset are replaced with the mean/median for that 5-minute interval
```{r}
steps_nafilledin <- activity
steps_nafilledin <- group_by(steps_nafilledin, interval) 
steps_nafilledin <- mutate(steps_nafilledin,steps=ifelse(is.na(steps), mean(steps, na.rm=TRUE), steps))


dim(steps_nafilledin)
class(steps_nafilledin)
str(steps_nafilledin)
summary(steps_nafilledin)
names(steps_nafilledin)
```
## Make a histogram of the total number of steps taken each day
```{r}
steps_nafilledin_days <- steps_nafilledin
steps_nafilledin_days <- group_by(steps_nafilledin_days, date) 
steps_nafilledin_days <- summarize(steps_nafilledin_days,steps=sum(steps))

dim(steps_nafilledin_days)
class(steps_nafilledin_days)
str(steps_nafilledin_days)
summary(steps_nafilledin_days)
names(steps_nafilledin_days)
head(steps_nafilledin_days)
tail(steps_nafilledin_days)
```

```{r, echo=FALSE}
ggplot(steps_nafilledin_days, aes(x=steps)) +
	geom_histogram(bins=20, colour="blue", fill="blue") +
	xlab("Steps")+ ylab("Count")+  
	labs(title= "Total numbers of Steps per day (missing data imputed)")

```
## Calculate and report the mean and median total number of steps taken per day of the new dataset.
#Do these values differ from the estimates from the first part of the assignment?
```{r}
steps_nafilledin$date <- ymd(steps_nafilledin$date)
total_steps_nafilledin <- tapply(steps_nafilledin$steps, steps_nafilledin$date, FUN = sum, na.rm = TRUE)


mean(total_steps_nafilledin)
median(total_steps_nafilledin)

mean(total_steps)==mean(total_steps_nafilledin)
median(total_steps)==median(total_steps_nafilledin)

summary(total_steps)
summary(total_steps_nafilledin)
```
## What is the impact of imputing missing data on the estimates of the total daily number of steps? 
## The estimates of the number of steps increased by 41, 3041, 370, 1416, 0, 0.
```{r, echo=FALSE}
summary(total_steps) - summary(total_steps_nafilledin)

par(mfrow=c(2,1))
hist(total_steps,col="blue")
hist(total_steps_nafilledin,col="blue")
```
#Part 3
## What are the differences in activity patterns between weekdays and weekends? 
## the dataset used in this part has filled-in missing values. 
```{r}
dayofweek <- function(date) {
    if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
        "weekend"
    } else {
        "weekday"
    }
}
steps_nafilledin$daytype <- as.factor(sapply(steps_nafilledin$date, dayofweek))

```
## A panel plot containing a time series plot (i.e. type = "l") of the 5-minute 
## interval (x-axis) and the average number of steps taken, averaged across all 
## weekday days or weekend days (y-axis). 
```{r}
par(mfrow = c(2, 1))
for (type in c("weekend", "weekday")) {
    steps_split <- aggregate(steps ~ interval, data = steps_nafilledin, subset = steps_nafilledin$daytype == 
        type, FUN = mean)
    plot(steps_split, type = "l", main = type)
}
```

